### 2016-11-03 section_discussion_subsection_Emission_Physics__.tex
added section_discussion_subsection_Emission_Physics__.tex

\section{discussion}

\subsection{Emission Physics and Burst Energetics}

\subsection{Connection to Overall FRB Population}

\subsection{Observing Strategies}