### 2017-04-25 repeater-vladata.tex
comments in as notes for all in but for jim

\documentclass[twocolumn]{aastex61}
%\documentclass{emulateapj}
\usepackage{hyperref} 
\usepackage{graphicx,natbib}
\citestyle{aa}
\usepackage[space]{grffile}
\usepackage{latexsym}
\usepackage{amsfonts,amsmath,amssymb}
\usepackage{url}
\usepackage[utf8]{inputenc}
\usepackage{fancyref}
\usepackage{multirow}
\hypersetup{colorlinks=false,pdfborder={0 0 0},}

\newcommand{\rf}{\emph{realfast}}
\newcommand{\frb}{FRB 121102}

\begin{document}

\title{The \frb\ Campaign: A Multi-Telescope Burst Detection and Implications for the FRB Population}
\shorttitle{\frb\ Observing Campaign}
\shortauthors{Law et al.}

\author[0000-0002-4119-9963]{C.~J.~Law}
\affiliation{Department of Astronomy and Radio Astronomy Lab, University of California, Berkeley, CA 94720, USA}

\author{M.~W.~Abruzzo}
\affiliation{Haverford College, 370 Lancaster Ave, Haverford, PA 19041, USA}

\author{C.~G.~Bassa}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}

\author{G.~C.~Bower}
\affiliation{Academia Sinica Institute of Astronomy and Astrophysics, 645 N. A'ohoku Place, Hilo, HI 96720, USA}

\author{S.~Burke-Spolaor}
\affiliation{National Radio Astronomy Observatory, Socorro, NM 87801, USA}
\affiliation{Department of Physics and Astronomy, West Virginia University, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505}

\author{B.~J.~Butler}
\affiliation{National Radio Astronomy Observatory, Socorro, NM 87801, USA}

\author{T.~Cantwell}
\affiliation{Jodrell Bank Centre for Astrophysics, Alan Turing Building, School of Physics \& Astronomy ,The University of Manchester, Oxford Road, Manchester M13 9PL, UK}

\author{S.~H.~Carey}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{S.~Chatterjee}
\affiliation{Cornell Center for Astrophysics and Planetary Science and Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}

\author{J.~M.~Cordes}
\affiliation{Cornell Center for Astrophysics and Planetary Science and Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}

\author{P.~Demorest}
\affiliation{National Radio Astronomy Observatory, Socorro, NM 87801, USA}

\author{R.~Fender}
\affiliation{Centre for Astrophysical Surveys, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK}

\author{K.~Grainge}
\affiliation{Jodrell Bank Centre for Astrophysics, Alan Turing Building, School of Physics \& Astronomy ,The University of Manchester, Oxford Road, Manchester M13 9PL, UK}

\author{J.~W.~T.~Hessels}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

\author{J.~Hickish}
\affiliation{Dept of Astronomy and Radio Astronomy Lab, University of California, Berkeley, CA 94720, USA}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{V.~M.~Kaspi}
\affiliation{Department of Physics and McGill Space Institute, McGill University, 3600 University St., Montreal, QC H3A 2T8, Canada}

\author{T.~J.~W.~Lazio}
\affiliation{Jet Propulsion Laboratory, California Institute of Technology, Pasadena, CA 91109, USA}

\author{M.~A.~McLaughlin}
\affiliation{Department of Physics and Astronomy, West Virginia University, Morgantown, WV 26506, USA}
\affiliation{Center for Gravitational Waves and Cosmology, West Virginia University, Chestnut Ridge Research Building, Morgantown, WV 26505}

\author{D.~Michilli}
\affiliation{ASTRON, Netherlands Institute for Radio Astronomy, Postbus 2, 7990 AA, Dwingeloo, The Netherlands}
\affiliation{Anton Pannekoek Institute for Astronomy, University of Amsterdam, Science Park 904, 1098 XH Amsterdam, The Netherlands}

\author{K.~Mooley}
\affiliation{Centre for Astrophysical Surveys, University of Oxford, Denys Wilkinson Building, Keble Road, Oxford OX1 3RH, UK}

\author{Y.~C.~Perrott}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{S.~M.~Ransom}
\affiliation{National Radio Astronomy Observatory, Charlottesville, VA 22903, USA}

\author{N.~Razavi-Ghods}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{M.~Rupen}
\affiliation{National Research Council of Canada, Herzberg Astronomy and Astrophysics, Dominion Radio Astrophysical Observatory, P.O. Box 248, Penticton, BC V2A 6J9, Canada}

\author{A.~Scaife}
\affiliation{Jodrell Bank Centre for Astrophysics, Alan Turing Building, School of Physics \& Astronomy ,The University of Manchester, Oxford Road, Manchester M13 9PL, UK}

\author{P.~Scott}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{P.~Scholz}
\affiliation{National Research Council of Canada, Herzberg Astronomy and Astrophysics, Dominion Radio Astrophysical Observatory, P.O. Box 248, Penticton, BC V2A 6J9, Canada}

\author{A.~Seymour}
\affiliation{Arecibo Observatory, HC3 Box 53995, Arecibo, PR 00612, USA}
\affiliation{Max-Planck-Institut f\"ur Radioastronomie, Auf dem H\"ugel 69, D-53121 Bonn, Germany}

\author{L.~G.~Spitler}
\affiliation{Max-Planck-Institut f\"ur Radioastronomie, Auf dem H\"ugel 69, D-53121 Bonn, Germany}

\author{S.~P.~Tendulkar}
\affiliation{Department of Physics and McGill Space Institute, McGill University, 3600 University St., Montreal, QC H3A 2T8, Canada}

\author{D.~Titterington}
\affiliation{Astrophysics Group, Cavendish Laboratory, 19 J. J. Thomson Avenue, Cambridge CB3 0HE, UK}

\author{R.~S.~Wharton}
\affiliation{Cornell Center for Astrophysics and Planetary Science and Department of Astronomy, Cornell University, Ithaca, NY 14853, USA}

\author{P.~K.~G.~Williams}
\affiliation{Harvard-Smithsonian Center for Astrophysics, Cambridge, MA, USA}

\begin{abstract}
We present results of the coordinated observing campaign that made the first arcsecond localization of a Fast Radio Burst, FRB 121102. Of the nine bursts detected by the Very Large Array at 3~GHz, three had simultaneous observing coverage at other observatories. Only one VLA burst was detected by another observatory, the Arecibo Observatory at 1.4~GHz. We use multi-observatory constraints and modeling of bursts seen only at 3~GHz to confirm earlier results showing that burst spectra are not well characterized by a power law. We characterize burst spectra with a $\sim500$~MHz envelope and apparent radio energy as high as $10^{40}$\ erg. We attribute significant changes in the apparent dispersion between bursts to frequency-dependent profiles that add a systematic random to the estimate of DM by up to 1\%. We use \frb\ as a prototype of the FRB class to estimate a volumetric birth rate of FRB sources $R_{\rm{FRB}} \approx 5\times10^{-5}$\ Mpc$^{-3}$\ yr$^{-1}$ for the cosmic volume out to redshift of 1. This rate is broadly consistent with models of FRBs from young magnetars or pulsars, if the typical FRB repeats on the order of $10^3$\ times in its life.

% ** be sure future high-burst rate paper from Arecibo doesn't rule this out **

\end{abstract}

\section{Introduction}

Fast Radio Bursts (FRBs) were discovered ten years ago with the detection of a millisecond-duration radio transient with an anomalously high dispersion measure \citep{2007Sci...318..777L}. The large measured DM implied that FRBs originated outside of our Galaxy, potentially at cosmological distances, and were orders of magnitude more luminous than any other millisecond radio transient \citep{2013Sci...341...53T}. There are now 21 FRBs publicly known \footnote{See \url{http://www.astronomy.swin.edu.au/pulsar/frbcat}.} with dispersion measures (DMs) as high as 1600 pc cm$^{-3}$ and intrinsic temporal widths of order milliseconds. Both their energetics and distance have inspired a wide variety of models and astrophysical applications \citep[e.g.,][]{2014ApJ...780L..33M, 2014ApJ...797...70K, 2015MNRAS.450L..71F, 2016MNRAS.458L..19C, 2016MNRAS.457..232C, 2016MNRAS.462L..16P}. However, that potential was limited by the lack of a definitive association of an FRB to an extragalactic host.

This paper is part of a series based on the first localization of an FRB and its unambiguous association to an extragalactic host \citep{LOC, OPT, EVN}. \frb, also known as the ``repeating FRB'', was first detected in November 2012 by the Arecibo Observatory \citep{2014ApJ...790..101S}. In mid 2015, new Arecibo observations revealed a series of bursts at a similar DM and sky position that rules out 
cataclymic models for this source \citep{2016Natur.531..202S}. The typical DM measured by early Arecibo observations was 559 pc cm$^{-3}$ \citep{2016arXiv160308880S}, although somewhat higher values have been seen in more recent observations (560.5 pc cm$^{-3}$; Hessels et al., in prep).

Beginning in August of 2015 (MJD 57623), we made the first of nine detections of \frb\ with the Karl G.\ Jansky Very Large Array \citep[VLA;][]{LOC} and localized it with a precision of 0.1\arcsec. Deep optical observing with the Gemini Observatory associated the FRB with a host galaxy at $z=0.193$\ \citep{OPT} and the European VLBI Network detected four more bursts to localize the source with a precision of 0.01\arcsec, four orders of magnitude better than any other FRB \citep[precision of $\sim40$\ pc in linear distance;][]{EVN}. The lookback and luminosity distances for \frb\ are 746 and 972 Mpc, respectively, assuming a concordance cosmology with parameters given by \citet{2016A&A...594A..13P}. At this distance, \frb\ is orders of magnitude more luminous than any other millisecond radio transient. If \frb\ is representative of all FRBs, then we should expect them to be useful as probes of the intergalactic medium and their host galaxies. Thus, the confirmation of a cosmological distance for \frb\ has begun to fulfil the promise implied by the first FRB detection.

Many new models of FRB origin have been been developed in response to the localization of \frb\ \citep{2017arXiv170104815K, 2017arXiv170102370M, 2017arXiv170104094Z, 2017arXiv170102492D, 2017arXiv170208644B, 2017arXiv170300393T}. The association of \frb\ with a compact, persistent radio source is consistent with bursts coming from a young magnetar that powers a luminous pulsar wind nebula \citep{2017arXiv170104815K}. At the same time, \frb\ is found in a low-metallicity dwarf galaxy \citep{OPT, 2017arXiv170102370M}, which is not a prediction of models that scale with stellar mass or star formation \citep{2017arXiv170400022N}. These galaxies are the preferred environment for long GRBs and Type-I superluminous supernovae (LGRB and SLSN-I, respectively), which have been suggested are signatures of the magnetar birth \citep{2008AJ....135.1136M, 2014ApJ...787..138L}.

%Also, it has not been demonstrated that \frb\ is representative of the overall FRB population. In fact, the repetition of its bursts is unique among all FRBs \citep{2015MNRAS.454..457P}, so it is natural to ask whether \frb\ is representative. An important first step is to demonstrate that the properties of \frb\ are consistent with the significant body of facts for the overall population \citep{2015MNRAS.451.3278M, 2016MPLA...3130013K}. The repeating nature of \frb\ provides us with a sample of bursts covering a range of spectral properties, temporal properties, and luminosities that can be used to test its connection to the larger FRB population.

This paper presents an analysis of the spectral properties of VLA bursts implied by simultaneous observing at Effelsberg, GBT, AMI-LA, and Arecibo, including the first simultaneous detection of an FRB at two observatories. If we assume that \frb\ is representative of all FRBs, we can use it to constrain the physical processes at play in the overall FRB population. The repetition of \frb\ also has strong implications for calculations of their rate of occurrence \citep{2016MNRAS.458L..89C} and comparison to other classes of transient, such as superluminous supernovae \citep{OPT}.
%More data is needed to determine if bursts trace intrinsic processes, such as coherent, pulsar-like emission \citep{2014PhRvD..89j3009K, 2014ApJ...785L..26L, 2016MNRAS.457..232C} or if extrinsic effects also play a role \citep{2015MNRAS.451.3278M, CORDES}.

In Section \ref{sec:obs}, we describe the multi-telescope observing campaign and a refined analysis of the nine VLA bursts. Section \ref{sec:res} presents the first spectrum of an FRB simultaneously detected at multiple telescopes and show that burst spectra cannot be modeled with a single spectral index. We then model the dynamic spectra to characterize the burst spectra, energies, and dispersion properties. Section \ref{sec:disc} discusses the properties of \frb\ bursts and its impact on inferences about the birth rate of FRB sources and strategies to find and/or localize new FRBs.

\section{Observations}
\label{sec:obs}

\begin{figure*}[t]
\begin{center}
\includegraphics[width=2\columnwidth]{timeline0}

\includegraphics[width=2\columnwidth]{timeline}
\caption{The top and bottom panels summarize the observations and detections of \frb\ in the 2015 and 2016 campaigns. Circles highlight observations that detected bursts from \frb\ with multiple burst detections indicated with a heavy line. The black dashed lines show the VLA burst detections with simultaneous coverage at other telescopes and the solid line shows the simultaneous burst detection. Circled observations with no line indicate observations on the same day that did not overlap. Symbols show days with observations with letters L, S, C, U referring to radio frequencies of 1.4, 3, 4.5, and 15~GHz, respectively.
\label{fig:sched}}
\end{center}
\end{figure*}

The data presented here were obtained from multiple programs and telescopes with a goal of interferometrically localizing \frb\ with the VLA. We coordinated observing between the VLA, Arecibo, Effelsberg, and AMI-LA telescopes, as shown in Figure \ref{fig:sched} and Table \ref{tab:obs}. Below, we summarize these observations, with a focus on those conducted simultaneously with VLA burst detections from \frb.

Computational (Jupyter) notebooks to reproduce the transient detection, localization, and analysis presented here can be found at \software{\url{https://github.com/caseyjlaw/FRB121102}}. Time cut-out visibility data and calibration products are available at \url{https://doi.org/10.7910/DVN/TLDKXG}. Original VLA visibility data are available under NRAO program codes 16A-459 and 16A-496 and can be downloaded at \url{http://archive.nrao.edu}.

\subsection{VLA}

The \frb\ observing campaign started in November 2015 with a 10~hr campaign ($\sim1$~hr per session) observed at 1.4~GHz in the compact D configuration. In April through May 2016, we conducted a 40~hr campaign ($\sim2$~hr per session) at 3~GHz in the C and CnB configurations in coordination with Arecibo \citep{2016arXiv160308880S}. We concluded with a 40~hr, coordinated campaign ($\sim2$~hr per session) from August through September 2016 in the B configuration and during the move to the most extended A configuration. In the late-2016 campaign, the first 34 hours of VLA observations were made at 3~GHz, while the last 6 hours were observed at 6~GHz. These observing session times are inclusive of calibration and overhead, which typically amounts to 15\% of the total observing time.

All VLA fast-sampled data were observed with 5~ms sampling, 256 channels, and dual-circular polarization \citep{2015ApJ...807...16L}. The channel frequency width was set to maximize sensitivity to the known DM of \frb\ with the largest total bandwidth. The total bandwidths at L (1.4~GHz), S (3~GHz), and C (6~GHz) bands were 256~MHz, 1024~MHz, and 2048~MHz, respectively. The 3~GHz data were recorded data in eight spectral windows with 32 channels each and had sensitivity of 5~mJy in 5~ms ($1\sigma$).

Observations in August and September were searched by a prototype version of \rf\ \citep[see \url{http://realfast.io};][]{2017AAS...22933002L}. \rf\ is a real-time, fast imaging transient search system. The current prototype runs on CPU-based hardware that is normally dedicated to the VLA correlator; for this experiment, it runs the transient search pipeline software called \software{\emph{rtpipe}\footnote{See \url{https://github.com/caseyjlaw/rtpipe}} and \citet{2015ApJ...807...16L}.}. Images were formed for each integration with DMs of 0, 546, 556.9, 560, and 565 pc cm$^{-3}$ and a time resampling grid of 5, 10, 20, 40, and 80 ms. This DM grid was chosen to maintain 90\% sensitivity to the nominal DM range of 540--570 pc cm$^{-3}$. Gain calibration was made from observations of J0555+3948 by an automated system (telcal), which uses phase-only calibration. A flux scale was calculated for each spectral window from an observation of 3C48 and applied to all burst spectra and has an accuracy of about 10\%.

Burst detections and localizations were made within hours of data being recorded. The transient search starts when data are recorded and proceeds slower than real-time, so we refer to it as ``quasi real-time''. For each trial DM, integration, and time scale, we form an image and calculate the S/N ratio for the peak pixel in the dirty image. We empirically identified S/N thresholds of 6.4 and 7.4 as useful to capture data quality statistics and candidates for inspection, respectively. The higher threshold is relatively unlikely to be triggered by thermal noise in this configuration, so \rf\ generates a more detailed (and computationally intensive) candidate visualization that includes an image and spectrum. All visibilities are recorded so detailed analysis, including improved calibration and localization, can be conducted offline.

\subsection{Arecibo}

During the late-2016 campaign, Arecibo observed with the L-wide receiver using the PUPPI pulsar backend. The observational frequency range was 1.15 to 1.73~GHz and frequency resolution was 1.5625~MHz, which has a typical sensitivity of 2~mJy in 2~ms ($1\sigma$). We recorded total full Stokes polarization intensity spectra with a time resolution of 10.24~$\mu$s. Each frequency channel was coherently dedispersed to 557~pc cm$^{-3}$, significantly reducing intra-channel dispersion smearing. The full width at half maximum (FWHM) beam size at band center is 3.3\arcmin.

In total, 11 Arecibo observations had some simultaneous coverage with the VLA. All of these observations were conducted after the VLA localization, so they were pointing directly at \frb. Three of those observations had simultaneous coverage of VLA bursts and one of those Arecibo observations detected the VLA burst. 
% During the first VLA burst with Arecibo coverage (MJD 57643), the PUPPI 1.4~GHz recording system failed so data were recorded at C band with the mock spectrometer. The C-band observations were recorded at frequencies from 4078 to 4245~MHz and had a sensitivity of 3.6~mJy in 2~ms ($1\sigma$). No detection was made in those Arecibo data. % levels set wrong. useless c-band ao data.
Overall, there were many more bursts detected at Arecibo than with the VLA, including some Arecibo bursts with simultaneous VLA upper limits. A more detailed analysis of those bursts will be presented elsewhere (Michilli et al, in prep).

\subsection{Effelsberg}

Effelsberg observations were conducted with the S60mm receiver and recorded total intensity spectra with the PFFTS backend in pulsar search mode. The observations had a time resolution of 65.5~$\mu$s and 128 frequency channels. The receiver has a system equivalent flux density of 18 Jy and a FWHM beam size of 2.4\arcmin\ at 4.85~GHz.

Five Effelsberg observations were made pointing at the known location of \frb\ and had some simultaneous coverage with the VLA. Two of these observations were simultaneous with VLA bursts. The nominal receiver bandwidth is from 4.6 to 5.1~GHz, but a configuration error reduced effective bandwidth to 100~MHz centered at 4.85~GHz for both of these observations during VLA bursts. The sensitivity in these observations was about 28 mJy in 2 ms ($1\sigma$), which is two times worse than the full-bandwidth value. No burst was detected in either observation.

\subsection{AMI-LA}

We observed \frb\ with the Arcminute MicroKelvin Imager Large Array (AMI-LA; Zwart et al. 2008) for 3 hours each on four epochs starting at MJDs 57638, 57643, 57645, 57646, 57648, and 57649. Observations were made with the new digital correlator having 4096 channels across a 5~GHz bandwidth between 13--18~GHz with a 1~s integration time. The phase calibrator, J0518+3306, was observed every 12 minutes for about 1.5 minutes. The AMI-LA data were binned to eight 0.625~GHz channels and processed (RFI excision and calibration) with a fully-automated pipeline, AMI-REDUCE \citep[e.g.,][]{2013MNRAS.429.3330P}. Daily measurements of 3C48 and 3C286 were used for the absolute flux calibration, which is good to about 10\%.

We inspected the calibrated visibilities, and did not find any signal above 30 mJy in the 1~s samples at and in the vicinity of the detected bursts. This corresponds to a sensitivity of 15~Jy for an assumed pulse width of 2~ms. Concatenating and imaging the 12 hours of calibrated data with the CASA tasks {\it concat} and {\it clean} also does not yield any significant detection at the FRB location. Although the statistical $3\sigma$\ upper limit is 60 $\mu$Jy, extended mJy-level sources in the field cause sidelobe confusion (the AMI-LA angular resolution is $\sim$30\arcsec), and the actual upper limit is larger. We introduced artificial point sources at the FRB location using the CASA {\it sm} tool, and found that these sources can be recovered at $3\sigma$\ as long as their peak flux densities are more than $\sim100\mu$Jy. Hence, we place an upper limit of $100\pm10 \mu$Jy on any quiescent or possible radio flaring on $\sim$days timescales from the FRB. This limit is similar to the flux density measured by the VLA at 12~GHz for the persistent source \citep{LOC}.

\section{Results}
\label{sec:res}

\subsection{Multi-Observatory Burst Spectrum}

Four of the VLA bursts had simultaneous observing coverage with either Arecibo, Effelsberg, or AMI-LA. Two of the VLA bursts had simultaneous observing coverage by all four observatories (Figure \ref{fig:sched}). Of these bursts, only one VLA burst (on MJD 57648) was simultaneously detected by another observatory (Arecibo).

% ** uncertainties in dm delay? **

Figure \ref{fig:sgram} shows a dynamic spectrum for this burst with both VLA and Arecibo data. The temporal resolution of the VLA data (5~ms) is much coarser than that of Arecibo, so data were resampled to the Arecibo temporal grid after correcting for barycentric and dispersion measure offsets. We assumed a DM of 560.5 pc cm$^{-3}$ (Hessels et al., in prep) to calculate a dispersion delay of 189 and 734~ms from infinite frequency to the top of the VLA and Arecibo bands, respectively. We used PRESTO \citep{2001PhDT.......123R} to calculate a relative barycentric time correction of 3.8~ms between the VLA and Arecibo at the time of observation.

The regridded dynamic spectrum has an apparent DM error that is evident both as a frequency-dependent time drift within the Arecibo observation and as an offset between the Arecibo and VLA bursts. Both of these drifts are consistent with an apparent DM of $\sim$565 pc cm$^{-3}$. More refined modeling of the VLA dynamic spectrum alone (\S \ref{sec:spec} and Table \ref{tab:spec}) are consistent with this higher DM value.

\begin{figure}[htb]
\begin{center}
 \includegraphics[width=\columnwidth]{aovla_spec.png}
 \caption{Composite dynamic specrum for the burst from \frb\ on MJD 57648 with data from the VLA and Arecibo observatories. The time axis is measured in milliseconds relative to MJD 57648.437788874842 with times corrected to the barycenter and infinite frequency (assuming DM$=560.5$\ pc $^{-3}$). The thick black line separates data from the VLA (2.5 to 3.5~GHz) and Arecibo (1.1 to 1.8~GHz) and flagged data are grey. The amplitude scale is normalized by assuming the Arecibo to VLA gain ratio is 10.
 \label{fig:sgram}}
\end{center}
\end{figure}

% ** errorbars on flux densities?

Figure \ref{fig:multi} shows coarse spectra built from integrated flux densities measured (or limited) for three bursts with multi-telescope observing coverage. The burst on MJD 57648 was detected with VLA and Arecibo with significances of 25$\sigma$\ and 39$\sigma$, which corresponds to peak flux densities in 5~ms of 111 mJy and 14 mJy at 3 and 1.4~GHz, respectively. Here we assume a typical pulse width of 2~ms to measure VLA and Arecibo integrated flux densities of 278 and 64 mJy at 3 and 1.4~GHz, respectively. Assuming a power law flux density model ($S_{\nu} \propto \nu^{\alpha}$), we find a spectral index $\alpha=1.9$. This is inconsistent with the spectral index limit implied by the Effelsberg nondetection at 4.5~GHz ($\alpha<-1.7$\ for a 2~ms burst and $5\sigma$ limit).

Overall, the bursts with simultaneous observing coverage are not well described by a power-law model. Two Arecibo nondetections and two Effelsberg nondetections of VLA bursts place mutually inconsistent lower and upper limits on the spectral index. The Arecibo and Effelsberg upper limits at 1.4 and 4.5~GHz, respectively, limit all burst spectral indices $\alpha_{1.4/3}>+4.8$ and $\alpha_{3/4.5}<-2.7$. The strictest limits on $\alpha_{1.4/3}$\ and $\alpha_{3/4.5}$\ are both derived from the burst on MJD 57649, which shows that a power law model is inappropriate even for a single burst.
% could note that these limits are understated by VLA integrated flux underestimate

\begin{figure}[htb]
\begin{center}
 \includegraphics[width=\columnwidth]{multispec.png}
 \caption{Broadband spectral measurements and limits for three bursts of \frb\ with simultaneous multi-telescope observing coverage (Arecibo at 1.4~GHz, VLA at 3~GHz, and Effelsberg at 4.5~GHz). Measurements are shown with dots and upper limits with triangles. Upper limits assume a pulse width of 2~ms and a $5\sigma$ detection threshold. Overlapping points are offset by tens of MHz in frequency for visualization.
 \label{fig:multi}}
\end{center}
\end{figure}

\subsection{VLA Bursts}
\subsubsection{Spectral Modeling}
\label{sec:spec}

This paper refines the analysis of the nine VLA radio burst spectra described in \citet{LOC} in a few ways. We use a better calibration scheme and have optimized the detection significance over a fine grid of DM ($\Delta \rm{DM}=1\ \rm{pc}\ \rm{cm}^{-3}$). After calibration and flagging, the visibility phases were rotated to the best-fit location \citep[RA, Dec $=$ 05h31m58.70s, +33d08m52.5s;][]{LOC} to extract a Stokes I spectrum that maximizes the image S/N for each burst.

Table \ref{tab:spec} summarizes the properties of all nine VLA bursts. Parameters such as integrated flux density, S/N, and Stokes V were measured from the burst properties integrated in frequency. Stokes V values estimated from the dual-circular feeds as $(RR-LL)/(RR+LL)\approx3\%$, which is comparble to the systematic error induced by the primary beam\footnote{See EVLA memo \#195 at \url{https://library.nrao.edu/public/memos/evla/EVLAM_195.pdf}.}. Given that systematic effects dominate the apparent circular polarization, this observation limits the fractional circular polarization to less than 3\%.

\begin{table*}
\caption{Properties of Bursts from \frb}
\centering
\begin{tabular}{cc|rrr|ccrrrr}
\multicolumn{2}{c|}{Time} & \multicolumn{3}{c|}{Observed properties} & \multicolumn{6}{c}{Modeled properties} \\ \hline
Calendar day & Burst time   & S/N & $S_{\rm{I, 5~ms}}$	& $S_{\rm{V}}$ 	& DM 			& W 			& $S_{\rm{I,peak}}$ & Center & FWHM & $E_{\rm{int}}$ \\
(2016)       &  (MJD)     &   (image)    & (mJy) 			& (mJy) 		    & (pc cm$^{-3})$ 	& (ms) 			& (mJy) 			& (GHz)  & (MHz) & ($10^{38}$\ erg)\\ \hline
23 Aug & 57623.74402686      & 38 		& 194			& +3				& 567$\pm2$ 		& 2.0$\pm0.2$	 		& 690 		& 2.8 		& 290 & 11 			\\
2 Sep & 57633.67986367      & 179 		& 1500			& --35 				& 568.2$\pm0.2$ 	& 2.05$\pm0.02$			& 3340 		& 3.2 		& 510 & 98				\\
2 Sep\tablenotemark{c} & 57633.69515938 & 15 & 69		& +2				& 562$^{+4}_{-6}$ 	& 2.5$^{+0.9}_{-0.6}$	& $>$430 	& $<$2.5	& $<$290 & 7  			\\
7 Sep & 57638.49937435      & 12 		& 55			& +5 				& 567$^{+7}_{-9}$ 	& 1.3$^{+1.4}_{-0.8}$	& 130 		& 3.1 		& 420 & 3  			\\
12 Sep\tablenotemark{a} & 57643.45730263  & 100 		& 508 & --5 		& 565.6$^{+0.6}_{-0.5}$ & 1.9$\pm0.1$		& 1170 		& 2.8 		& 510 & 34 			\\
14 Sep\tablenotemark{a} & 57645.42958602   & 13 		& 64 & +3			& 563$^{+5}_{-4}$ 	& 1.1$\pm0.7$		 	& 170 		& 2.8 		& 380 & 4  			\\
15 Sep\tablenotemark{c} & 57646.46600650 & 20 & 87		& +1				& 569$\pm5$ 		& 2.5$^{+0.9}_{-1.4}$	& $>$420 	& $<$2.5 	& $<$430 & 10 			\\	
17 Sep\tablenotemark{a,b} & 57648.43691490 & 25 & 111	& +9				& 564$\pm2$ 		& 1.4$^{+0.3}_{-0.4}$	& 260 		& 2.8 		& 470 & 7  			\\
18 Sep\tablenotemark{a} & 57649.45175697   & 36 		& 167 & +1			& 567$\pm2$ 		& 2.1$\pm0.5$		 	& 290 		& 3.0 		& 690 & 12 			\\ \hline
\end{tabular}
\tablecomments{$S_{\rm{V}}$ is the measured circular polarization of the burst, which dominated by systematic effects. All error ranges represent 68\% confidence intervals.}
\tablenotetext{a}{Simultaneously observed by either Arecibo, Effelsberg, or AMI-LA.}
\tablenotetext{b}{Detected simultaneously with Arecibo.}
\tablenotetext{c}{Best-fit Gaussian is not centered in 3~GHz band, so spectral parameters are limits.}
\label{tab:spec}
\end{table*}

Figure \ref{fig:spec} shows that the Stokes I spectra are generally characterized by a broad, Gaussian shape with inter-channel modulation as high as 100\%. Diffractive scintillation from the Milky Way has a typical bandwidth of 7~MHz along this line of sight \citep{2002astro.ph..7156C}, which is similar to the channel size of 4~MHz. The burst energy is calculated from the flux density in 5~ms and integrating over the optimal Gaussian spectral model.

We used a Markov chain Monte Carlo (MCMC) method to model the dynamic radio spectra (time versus frequency). We used the Goodman and Weare affine invariant sampler \citep{goodman2010ensemble} as implemented in \emph{emcee} \citep{2013PASP..125..306F}. The dynamic spectra were modeled with a broad Gaussian envelope (amplitude, center, width), a time delay according to the $\nu^2$\ dispersion law, and a square temporal width (W). The inter-channel amplitude modulation was ignored in the model.

We use a log likelihood function $\ln \mathcal{L}=(-1/2)\ \sum_i (S_i - S_{i, \rm{model}})^2/\sigma_s^2$, where $S_i$\ refers to the measured and modeled flux density in a single pixel of the dynamic spectrum and $\sigma_s\approx 70\ \mu$Jy is the measured off-burst noise per 5~ms integration and 4~MHz channel. A flat prior was used over all ranges with valid data with the requirement that the integrated signal must have a detection significance higher than 8$\sigma$. The 6-dimensional models were sampled with 100 chains taking 700 steps; we ignored the first 200 steps to properly sample the posterior distributions.

The parameters for the best model of the dynamic spectrum are given in Table \ref{tab:spec} and the resulting Gaussian model overlaid on Figure \ref{fig:spec}. The typical burst spectrum has a FWHM of 500~MHz. All but two of the best-fit Gaussians are centered inside the 3~GHz band and some appear contained by the 1~GHz wide band. This is consistent with previous detections of \frb\ by Arecibo, which showed quasi-broadband structure \citep{2016arXiv160308880S} and large variation in the implied spectral index \citep{2014ApJ...790..101S}.

% ** quantify how much of model shape is actually inside the band **

% ** confusion as to whether inter-channel modulation is signal or noise ** jim also to do noise analysis **

% ** spectra plot modifications: **
% y-axis scale not shown! 
% add mjd to each plot
% Need to make sure impression differs of noise and inter-channel modulation

\begin{figure*}[ht]
\begin{center}
 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57623.png}
  \includegraphics[width=0.5\columnwidth]{spec_57633_scan7.png}
 \end{minipage}

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57633_scan13.png}
  \includegraphics[width=0.5\columnwidth]{spec_57638.png}
 \end{minipage}

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57643.png}
  \includegraphics[width=0.5\columnwidth]{spec_57645.png}
 \end{minipage}

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57646.png}
  \includegraphics[width=0.5\columnwidth]{spec_57648.png}
 \end{minipage}

 \begin{minipage}{2\columnwidth}
  \includegraphics[width=0.5\columnwidth]{spec_57649.png}
 \end{minipage}
\caption{Spectra and best-fit model for nine bursts seen by the VLA from 2.5 to 3.5~GHz in the peak, dedispersed 5~ms integration. Starting at the top left (moving right and then down), they correspond to bursts on MJDs 57623, 57633.68, 57633.70, 57638, 57643, 57645, 57646, 57648, and 57649. The solid line is a best-fit Gaussian model found through modeling.
\label{fig:spec}}
\end{center}
\end{figure*}

This analysis differs from traditional analysis in that it simultaneously models the spectral and temporal evolution of the burst. A typical \frb\ burst takes $\sim180$~ms or 36 integrations to cross the observing band from 2.5 to 3.5~GHz. The effect of a finite pulse width is visible as the burst moves from one integration to the next, even for widths narrower than the integration time of 5~ms. Table \ref{tab:spec} shows that the typical burst width is measured as $2\pm0.5$~ms (68\% confidence interval); the inter-channel dispersion smearing ranges from 0.4 to 1.1~ms across this band. The brightest burst (on 57633.68) is modeled with a temporal width of $2.05\pm0.02$~ms and its arrival time is measured with a precision of $\sim50\mu$s. Note that these errors are only accurate to the degree that the model represents the data. One potential bias in the model is that we do not model scintillation effects.

% ** fix plot to show "W" instead of "dint"

\begin{figure*}[htb]
\begin{center}
\includegraphics[width=2\columnwidth]{corner57646}
\caption{A scatterplot matrix that shows the correlation of every pair of parameters in the MCMC run for burst 57646. The equivalent plot for the other bursts show normally-distributed samples, but this burst shows significant structure in the samples. The parameters ``Amp'', ``Center'', and ``$\sigma_{ch}$'' refer to the Gaussian spectral shape, while ``DM'', ``int'', and ``W'' refer to the temporal shape. The parameters are given in units of channels (4 MHz) and integrations (5 ms).
\label{fig:corner}}
\end{center}
\end{figure*}

While most bursts are modeled with well-defined parameter probability distributions, one burst is an outlier. Figure \ref{fig:corner} shows the scatter plot matrix \citep{corner} for the MCMC run on burst 57646. Two clusters of samples are identified for this burst: one narrow, low-DM and one wide, high-DM. Assuming that the model is appropriate, this implies that the burst can be decomposed into two subbursts.

\subsubsection{Dispersion}

The initial VLA detections were made with a matched-filter approach that allowed inter-DM sensitivity losses up to roughly 10\% \citep[$\Delta \rm{DM}=10\ \rm{pc}\ \rm{cm}^{-3}$][]{2003ApJ...596.1142C}. 
%In optimizing detection significance over a fine DM grid, we find optimal DM values range from 552 to 572 pc cm$^{-3}$. However, the uncertainty in the peak DM measurement is defined by the signal-to-noise of the burst and how this changes with DM.
The MCMC modeling is more sophisticated in that it considers a wider range of parameters and models both data and observing effects (frequency-time pixel boundaries) simultaneously.

Figure \ref{fig:dmmodel} compares the 68\% confidence intervals on the DM for all nine VLA bursts against detection time (top panel) and the modeled burst temporal width (bottom panel). The error weighted mean burst DM is 567.8$\pm$0.1 pc cm$^{-3}$, significantly larger than the long-term average of 560.5 pc cm$^{-3}$ seen by Arecibo during this campaign. However, we note that the 95\% confidence intervals in DM for multiple bursts are not consistent with this mean nor each other, which suggests that some burst-specific property can bias the measured DM. The variation in DM observed by the VLA at 3~GHz is similar to that reported for Arecibo observations of \frb\ in \citet{2016arXiv160308880S}.
%As noted before, this is consistent with the idea that the DM measured at low temporal resolution (e.g., by the VLA or by single-dish observatories without coherent dedispersion) is determined by extrinsic (an electron column) dispersion and some spectrotemporal structure that changes between bursts.

The bottom panel of Figure \ref{fig:dmmodel} compares the DM to the modeled temporal width of the bursts. There is a weak correlation between burst width and apparent DM. A change in width of $\sim1$~ms correlates with a change in apparent DM of approximately $5$~pc cm$^{-3}$.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=\columnwidth]{burst_dmdt}
\caption{(Top) Burst time versus DM for nine 3~GHz bursts detected by the VLA. The DM scale for both panels is shown at the bottom. The 68\% confidence interval on DM is shown with a bar and the dahsed line shows the best-fit DM$=560.5$\ pc cm$^{-3}$ inferred from Arecibo observations at 1.4~GHz (Hessels et al, in prep). (Bottom) Temporal width versus burst DM for the same nine VLA bursts.
\label{fig:dmmodel}}
\end{center}
\end{figure}

\subsection{Temporal Statistics}
\label{sec:temp}

The burst rate for \frb\ varied dramatically throughout the 2016 observing campaign (Figure \ref{fig:sched}). In the early-2016 campaign, we observed for 30 hours at 3~GHz and no bursts were detected. In the late-2016 observing campaign, we observed for 27 hours at 3~GHz and detected nine bursts. Overall, the data quality is uniform and high, so the inhomogeneous burst distribution shows that the burst rate was not uniform.

Assuming that the burst detection probability follows a Poisson distribution, the mean VLA, 3~GHz burst rate is $R=0.16\pm0.05\ \rm{hr}^{-1}$ above a fluence of 0.2 Jy ms. The nondetection in the first half of 3~GHz observing limits the FRB rate to $R<0.1$\ hour$^{-1}$\ (95\% confidence limit). The 3~GHz burst rate was much higher during the late-2016 campaign, $R=0.3\pm0.1$\ hour$^{-1}$.

\subsection{Energy and Brightness Distribution}
\label{sec:disn}

Knowing the source distance, we calculate the burst energy by integrating flux in time and frequency (Table \ref{tab:spec}). Some VLA burst spectra seem to be contained by the 2.5 to 3.5~GHz band and most of them seem to have Gaussian envelopes that are well modeled by the emission within that band. Assuming that the Gaussian shape defines the full emission window, the mean 3~GHz flux density can be converted to a total energy with no further assumptions about the burst spectral properties.

Figure \ref{fig:ed} shows the \frb\ burst energy cumulative distribution as seen by the VLA and calculated from prior observations by Arecibo and the GBT \citep{2016Natur.531..202S,2016arXiv160308880S}. The latter two energy distributions are scaled from the fluence by assuming that all burst energy is included by the observation. This likely underestimates the burst energy for some bursts by a factor of a few, but should not affect the slope of the distribution.

Broadly, the range of energies shown in Figure \ref{fig:ed} is defined by the sensitivity of the telescope and the duration of the observing campaign. The VLA distribution shown represents the 27-hour, late-2016 campaign, which is longer than the single-dish campaigns presented here. However, since the VLA observations have a relatively coarse time resolution, it is not as sensitive as Arecibo and the GBT, so no bursts with energiest less than $3\times10^{38}$~erg were detected. The rate upper limit (95\% confidence) from the early-2016 VLA campaign shows that even identical observing campaigns have different detection rates.

We modeled the differential energy distribution, $dN/d\log{E}$, again using a Poisson detection probability with a rate function $\lambda = A E^{\alpha}$. Rather than trying to estimate a completeness limit for each energy distribution, we assume an effective detection limit of 0.9 times the weakest burst detected; the best-fit slope is weakly dependent on this, but general conclusions are robust. We directly sampled the likelihood distribution to estimate a best slope of $\alpha_{\rm{VLA}}=-0.7^{+0.3}_{-0.4}$, $\alpha_{\rm{Arecibo}}=-0.8^{+0.3}_{-0.5}$, and $\alpha_{\rm{GBT}}=-0.8^{+0.4}_{-0.5}$ (68\% confidence interval). Expressed as a power law function in $dN/dE$, the slope index is approximately --1.7.

\begin{figure}[htb]
\begin{center}
\includegraphics[width=\columnwidth]{energy_disn}
\caption{The cumulative burst energy distributions for the VLA, Arecibo, and GBT bursts are shown with dots connected by solid, dashed, and dotted lines, respectively. The Arecibo distribution is derived from the 11 1.4~GHz bursts reported in \citet{2016Natur.531..202S} and the GBT distribution is derived from the 5 1.4~GHz bursts reported in \citet{2016arXiv160308880S}. An upper limit from the VLA nondetection in early 2016 is shown as a triangle. \label{fig:ed}}
\end{center}
\end{figure}

\section{discussion}
\label{sec:disc}
\subsection{Burst Spectra}

We present the first simultaneous detection of an FRB with multiple telescopes and over frequencies from 1.2 to 3.5~GHz. The flux density of the Arecibo burst detected on MJD 57648 is an order of magntude less than that seen by the VLA. At the same time, three other bursts from \frb\ had similar observing coverage but were not detected simultaneously. This is consistent with the spectral structure observed within the 3~GHz VLA band, which is typically limited to a Gaussian envelope with width of roughly 500~MHz. This confirms earlier results at 1.4~GHz \citep{2016Natur.531..202S,2016arXiv160308880S} with a wider bandwidth and extends the presence of this phenomenon to 3~GHz.

If \frb\ burst spectra are typical of the larger FRB population, then population models need to be generalized beyond the assumption of a spectral power law. Most obviously, \frb\ implies that future multi-telescope searches for bursts are unlikely to simultaneously detect bursts in different bands \citep[c.f.][]{1999ApJ...517..460S}. Second, since burst spectra have limited bands, the burst detection rate at one frequency may not represent that at another. Rate estimates for the FRB population will need to explicitly account for the frequency-dependent rate. Finally, we note that the odds of detecting a burst will improve with bandwidth beyond the gain in sensitivity, since bandwidths wider than the FRB characteristic width are more likley to cover the burst envelope. However, FRB search algorithms will need to be modified to search for bursts with spectral width less than the full bandwidth.

After correcting for barycentric and typical dispersion delays, the VLA/Arecibo burst spectrum shows a residual temporal drift. This drift can be interpreted as an excess DM, but contemporaneous observations at 1.4~GHz measured a significantly lower DM (560.5 pc cm$^{-3}$; Hessels et al, in prep). This excess DM is comparable to the burst-to-burst variation in DM within the whole sample of VLA 3~GHz bursts. This variation had been noted in earlier 1.4~GHz Arecibo observations of \frb\ \citep{2016arXiv160308880S}, but the VLA bursts demonstrate that this burst-specific DM is seen up to 3~GHz.

One interpretation for the burst-dependent DM is that the bursts have a frequency-dependent pulse shape. Pulsar pulse shapes evolve with frequency on GHz frequency scales, presumably due to changes in the beam shape \citep{1988MNRAS.234..477L}. The burst-dependent DM changes are on the order of 1\% of the total DM (equivalent to a delay rate up to 2~ms GHz$^{-1}$), which is much larger than typically observed \citep{2017MNRAS.466.3706L}. The systematic offset of the VLA 3~GHz DM from that observed by Arecibo suggests that any such burst frequency evolution must be larger at higher frequencies. Our modeling of the VLA burst dynamic spectra show a weak correlation between larger apparent DM and pulse width. This could be a signature of unresolved spectrotemporal structure in the VLA burst spectra that biases the measured DM.

% ** expand on implication of spectral envelope in scintillation models **

The burst-specific DM structure could be intrinsic to the emission mechanism or induced during propagation. Radio waves can be modified in a variety of ways (e.g., scintillation, scattering) and the duration of the emission in the source frame is not known \citep{2016arXiv160505890C}. \citet{CORDES} describe a model where plasma lensing near the source of an FRB can magnify radio emission by orders of magnitude. That model argues that the simultaneous appearance of a burst from 1.2 to 3.5~GHz requires a lens with the focal frequency be $\gtrsim 3.5$~GHz. This gives a constraint on the combination 
$$\frac{\rm DM_0}{a_{\rm AU}^2} \frac{d_{sl}d_{lo}/d_{so}}{1\, \rm kpc} \gtrsim (3.5 / 39.1)^2$$
where DM$_0$, $a$, and $d_{sl, lo, so}$\ are the electron column density of the lens, size of the lens, and lensing distances, respectively. A lens placed at $d_{sl}\approx1$~kpc is consistent with typical properties of the interstellar medium. Lenses that are influenced by the source would be much nearer to the source $d_{\sl} \ll 1$~kpc, requiring large values of $\rm{DM_0}/a_{\rm AU}^2$.

\subsection{Energy Distribution}

% ** discuss beaming **

Our refined analysis shows that \frb\ has even higher isotropic energies than previously reported for \frb, reaching energies $E_{\rm{max}}\approx10^{40}$\ erg. \frb\ has isotropic energies orders of magnitude higher than the strongest Crab giant pulses ($\sim10^{35}$~erg) and comparable brightness temperatures \citep[$T_{b,\rm{Crab}}\sim10^{41}$~K versus $T_{b, \rm{FRB 121102}}\sim10^{38}$~K;][]{2003Natur.422..141H,2014PhRvD..89j3009K}. 
%While this emission is coherent and strongly beamed \citep{2016Natur.531..202S, WEIRD}, the energy scale is still larger than any known Galactic radio transient and requires either a new process or dramatic scaling of known emission processes \citep{2016MNRAS.462..941L, 2016MNRAS.457..232C}.

In \S \ref{sec:disn} we compare the energy distribution of the 9 VLA bursts to those reported earlier by Arecibo and the GBT. All three energy distributions can be characterized as a power law in $dN/dE$ with slope $\sim-1.7$. This slope is seen even though the burst rate varies by almost an order of magnitude between observations and frequencies cover both 1.4 and 3~GHz bands. This suggests that the slope is related to the underlying physical process, rather than the burst detection rate at any given time. The value of the energy distribution slope is similar to that observed in magnetars \citep{2000ApJ...532L.121G, 2011ApJ...739...94S}. This similarity has been used to argue that bursts are driven by self-regulated critical phenomena \citep[slope$=-5/3$;][]{2011SoPh..274...99A}.

\subsection{Flux Distribution}

While the surprising brightness of new FRB discoveries was once seen as a cause for concern (i.e., a sign of terrestrial interference), we may be able to use this to infer properties of the overall population \citep{2007Sci...318..777L, 2016arXiv161105758R}. When modeling the flux distribution as a power law, an isotropic distribution of sources will have a power law index, $\alpha$, of --1.5. Prior to the measurement of the first GRB redshift, deviations from this Euclidean distribution were used to infer their cosmological distribution \citep[e.g., the $V/V_{\rm{max}}$\ test;][]{1992ApJ...388L..45M, 1995ApJ...453...25F}. Multiple studies have suggested that FRBs have a sub-Euclidean flux distribution \citep[$-0.5<\alpha<-0.9$;][]{2016ApJ...830...75V, 2016arXiv160206099L, 2016arXiv161100458L}. Others have noted that the flux distribution is not well modeled as a power law, such that $\alpha$\ is effectively sensitivity (and telescope) dependent \citep{2016MNRAS.461..984O, 2017arXiv170208040C}.

We now know that \frb\ is detectable with the VLA out to $z=0.7$, which suggests that cosmological effects are even more important for understanding properites of the FRB population. \citet{2017arXiv170208040C} demonstrate how redshift space distortions, time dilation, and spectral index (``k-correction'') effects can flatten the flux distribution. To this complex scene, we add the fact that \frb\ burst spectra are not defined by a spectral index, which suggests that k-corrections are likely to be very difficult to calculate in practice. If the FRB burst rate is higher at low (rest) radio frequencies, then high-redshift FRBs will have an lower (redshifted) rate that flattens the $dN/dF$\ distribution in a manner similar to a k-correction. However, propagation effects are expected to suppress the FRB rate at sub-GHz frequencies \citep{2017MNRAS.465.2286R, 2017arXiv170107457C} as well. The VLA and Arecibo observed intensively in a coordinated fashion during the late-2016 campaign and a comparison of their relative rates will be presented elsewhere.

%While the \frb\ burst energy distribution is consistent with the overall FRB population, this simulation ignores many observational biases (scattering, observing coverage) and the effects of galaxy evolution, which is likely significant over the range of redshifts to which \frb\ is detectable \citep[e.g.,][]{2010ApJ...709..644I}. The burst energies are estimated from observations from 2.5 to 3.5~GHz, which may be a biased selection of all bursts or may not include all burst emission. If the simulation is revealing a connection between \frb\ and the broader FRB population, it does not explain the physical mechanism that drives it. The burst energy distribution of the population could be dominated by intrinsic or extrinsic effects \citep{2015MNRAS.451.3278M,2016Natur.531..202S,CORDES}. More FRB detections will improve the quality of this inference as sample variance drops, while more burst detetions from \frb\ will improve the measurement of the intrinsic energy distribution.

\subsection{Repetition}

The chance of detecting a burst from \frb\ with the VLA changes substantially on day--month timescales. A major outstanding question is whether this time-variable rate is driven by an intrinsic process \citep{2016ApJ...826..226K} or extrinsic (propagation) effects in the host galaxy or intergalactic medium \citep{CORDES}. However, the temporal distribution of VLA detections alone shows that the \frb\ burst rate is significantly correlated on short timescales. That is, detecting a burst predicts that you are more likely to find another burst soon thereafter and not detecting a burst predicts you will likely not detect another burst soon thereafter.

The burst time distribution is consistent with a ``red spectrum'' \citep{2016MNRAS.458L..89C}, although the observations presented here are only sensitive to timescales of days (time between observations) and months (time between campaigns). As has been noted elsewhere, this statistical property would weaken previous constraints on repetition \citep{2015MNRAS.454..457P,2015ApJ...807...16L}. If this statistical property describes other FRBs, then wide, shallow surveys are the preferred strategy for blind detection of FRBs. Also, since ``bursts predict bursts'', many short observations are more likely to detect an FRB than a single long observation of the same total length and all new FRB detections should be immediately and intensively followed up.

\subsection{Volumetric Rate of FRB Sources}

%The calculation of a volumetric rate has been used to study classes of optical and high-energy transient \citep{2006ARA&A..44..507W}. 
\citet{2013Sci...341...53T} estimated an isotropic rate of FRB events of $10^{-3}$\ galaxy$^{-1}$\ yr$^{-1}$ by calculating the number of Milky Way-type galaxies out to a distance implied by assming all extragalactic DM originates in the IGM \citep[$z\approx0.9$\ for DM$\approx z\times900\ \rm{pc}\ \rm{cm}^{-3}$;][]{2003ApJ...598L..79I,2004MNRAS.348..999I}. This rate is high \citep[comparable to the rate of core-collapse supernovae;][]{2006Natur.439...45D}, which has been used to argue that FRBs are likely not associated with other classes of transient, such as LGRBs or rare subclasses of SNe \citep{2006ARA&A..44..507W}. However, since this rate depends on telescope sensitivity, it is not appropriate to compare to other estimates based on source counts.

Here, we reevaluate the FRB volumetric rate by assuming \frb\ is a prototype of the class and recasting it as a volumetric rate of FRB sources (i.e., a birth rate). Crudely, the volumetric rate of FRB sources can be defined as $R_{\rm{FRB}} = R_p /(N_r \Omega_b V (z))$, where $R_p$\ is the projected FRB rate, $N_r$\ is the number of bursts per source in a typical lifetime, $\Omega_b$\ is the beaming fraction, and $V(z)$\ is the comoving volume out to redshift $z$. The newest estimates of the projected FRB rate are $2\times10^3\ \rm{sky}^{-1} \rm{day}^{-1}$\ at high Galactic latitudes and flux densities brighter than 1~Jy~ms \citep{2016arXiv161100458L, 2016MNRAS.460L..30C, 2016MNRAS.455.2207R}. There is relatively little constraint on the beaming factor, but Galactic pulsars have beaming solid angle fractions on the order of 10\% \citep{1998MNRAS.298..625T}.

\frb\ has an absolute energy scale that can be used to estimate a horizon scale for the typical FRB survey sensitivity. At the typical Parkes survey parameters \citep[e.g.,][]{2016MNRAS.460L..30C}, a survey with a fluence limit of 1~Jy~ms can detect \frb\ at $z\approx1$. The distance inferred from the largest DMs observed \citep[$\sim1500$\ pc cm$^{-3}$;][]{2016MNRAS.460L..30C} is somewhat higher, but that likely overestimates distance since DM is expected to have significant contributions from the host galaxy and intervening galaxies \citep{OPT, 2014ApJ...780L..33M}.
%For reference, we note that cosmic star formation rate peaks between redshift of 2 and 3\citep{2014ARA&A..52..415M}.

Using an assumed horizon redshift of 1, we estimate a rate of $R_{\rm{FRB}} \approx 5\times10^{-5} N_r^{-1} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$. This calculation ignores time dilation, which underestimates the burst rate by a factor of 2 at the redshift horizon. The narrow burst spectral structure seen for \frb\ also suggests that the FRB rate is understated by a factor of a few by ignoring bursts that fall outside the observing band. Finally, we note that this estimate assumes that Galactic latitude and propagation effects do not significantly affect the detectability of FRBs, although scintillation and scattering likely play significant roles \citep{2015MNRAS.451.3278M, CORDES}.

Despite its many assumptions, this volumetric rate is more appropriate for comparison to rates used for other classes of transient. The LGRB and SLSN-I rates are $10^{-7}$\ and $10^{-8}$\ Mpc$^{-3}$\ yr$^{-1}$, respectively \citep{2007ApJ...657L..73G,2012Sci...337..927G}. Comparing the FRB rate to these rates implies that FRB-emitting sources can be generated in LGRBs and SLSN-I if FRBs repeat $5\times10^2$\ and $5\times10^3$\ times\footnote{\citet{2017arXiv170400022N} provide a similar estimate of FRB birth rate and reach a similar conclusion.}, respectively (assuming burst energies of $\sim10^{40}$\ erg). \citet{2017arXiv170102370M} use the young magnetar model for FRBs to calculate the maximum number of bursts that can be powered its magnetic field as $\approx300$. Assuming this model for FRBs, LGRBs, and SLSN-I, the rate comparison suggest that the LGRB rate is most consistent with the FRB rate, although there is significant room for adjustment in the models and rate estimate.

%\subsection{Naming Convention}
%The consistency of \frb\ properties (e.g., energy distribution) with the overall population suggests other FRBs likely repeat. If so, the current naming convention for FRBs (analogous to cataclysmic supernovae), will likely become uninformative, especially as large FRB survey projects come online \citep{2014SPIE.9145E..22B}. If all FRBs repeat, then a more useful convention would be based on coordinates. However, there are already two FRBs that have consistent celestial positions and different DMs. Therefore, we suggest an FRB naming convention as Jhhmm+ddDMmmmm.

\section{Conclusions}

The recent precision localization of \frb\ has helped identify a host galaxy, measure its distance, and establish it as a member of a truly new class of astrophysical source. With its cosmological distance firmly established, \frb\ now serves as a new kind of standard by which FRBs are defined. That fact, combined with a rich data set of many bursts, allows us to apply this source to more general questions about the FRB population.

We presented the first multi-telescope detection (Arecibo and VLA) of an FRB. By detecting this burst from 1.2 to 3.5~GHz, we have demonstrated that some bursts have broad spectral structure. However, the energy of that burst was dominated by the VLA 3~GHz observing band and three other VLA bursts are undetected by Arecibo. This demonstrates the burst spectra are poorly described by a power law with a single spectral index. Independently, we modelled dynamic spectra within the VLA 3~GHz band and show that most bursts can be characterized by a Gaussian envelope of width $\sim500$\ MHz. This modeling also shows that the apparent DM changes from burst to burst and that DM is biased above the long-term average measured at 1.4~GHz. The nature of this DM change is not known, but could be explained by strong frequency-dependent profile evolution or unresolved spectrotemporal in the bursts.

With a characteristic burst spectrum, we can estimate total (apparent) radio energy per burst. The cumulative energy distribution is characterized by a power law in $dN/dE$ with slope of $\sim-1.7$. The amplitude of this power law changes significantly between observational campaigns. The stochastic nature of the burst rate suggests that past constraints on FRB repetition are weaker than previously inferred. The relatively narrow spectral structure, flat energy distribution, and variable burst rate of \frb\ suggests that repeated observations, wide bandwidth, and large instantaneous field of view all improve the odds of FRB discovery.

Assuming that \frb\ is representative of the FRB population, we calculate a volumetric birth rate of FRB sources that does not depend on telescope sensitivity. We estimate a volumetric rate of FRB sources of $R_{\rm{FRB}} \approx 5\times10^{-5} (0.1/\Omega_b)$\ Mpc$^{-3}$\ yr$^{-1}$, which is appropriate for \frb-like burst energies of $10^{40}$\ erg that are detectable out to $z=1$. This rate is broadly consistent with models of FRBs from young magnetars or pulsars if the typical FRB repeats on the order of $10^3$\ times over its lifetime.

New, arcsecond-scale localizations will be critical to refining the picture presented here and constraining models of FRB origin. \frb\ was localized within hours by a prototype version of \rf\ and an expanded \rf\ system is now under construction. This platform will search a TB/hour data stream in real time in parallel with ongoing VLA observations, potentially detecting and localizing multiple FRBs per year.

\appendix

% ** add mjd colum **
% ** use footnotes instead of last two cols? **

\section{Table of Observations}

\startlongtable
\begin{deluxetable*}{lcccccc}
\small
\tablecaption{Table of Observations \label{tab:obs}}
%\centering
%\begin{tabular}{lcccccc}
\startdata
Day & Start time & Duration & Band & VLA Detections & Other Coverage & Other Detections \\
(yyyy-mm-dd; UT) & (hh:mm:ss; UT) & (minutes) & & & &  \\ \hline
2015-11-25 & 3:25:29 & 60 & 1.4~GHz & 0 &  & \\
2015-12-01 & 5:31:31 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-03 & 2:53:57 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-05 & 4:45:44 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-07 & 4:37:50 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-09 & 9:29:25 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-11 & 9:22:27 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-13 & 9:13:7 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-15 & 9:6:16 & 60 & 1.4~GHz & 0 &  &  \\
2015-12-17 & 8:57:51 & 60 & 1.4~GHz & 0 &  &  \\
2016-04-25 & 15:39:33 & 42 & 3~GHz & 0 &  &  \\
2016-04-26 & 21:49:35 & 15 & 3~GHz & 0 &  &  \\
2016-04-30 & 19:14:46 & 60 & 3~GHz & 0 &  &  \\
2016-05-01 & 18:8:21 & 60 & 3~GHz & 0 &  &  \\
2016-05-03 & 22:53:26 & 60 & 3~GHz & 0 &  &  \\
2016-05-04 & 22:50:30 & 60 & 3~GHz & 0 &  &  \\
2016-05-06 & 18:25:12 & 60 & 3~GHz & 0 &  &  \\
2016-05-07 & 19:34:13 & 60 & 3~GHz & 0 &  &  \\
2016-05-08 & 19:30:20 & 60 & 3~GHz & 0 &  &  \\
2016-05-13 & 17:12:48 & 135 & 3~GHz & 0 &  &  \\
2016-05-14 & 17:14:11 & 135 & 3~GHz & 0 &  &  \\
2016-05-15 & 17:15:26 & 135 & 3~GHz & 0 &  &  \\
2016-05-16 & 16:59:52 & 135 & 3~GHz & 0 &  &  \\
2016-05-20 & 16:45:46 & 135 & 3~GHz & 0 &  &  \\
2016-05-21 & 16:44:6 & 135 & 3~GHz & 0 &  &  \\
2016-05-22 & 16:43:53 & 135 & 3~GHz & 0 &  &  \\
2016-05-23 & 16:44:11 & 135 & 3~GHz & 0 &  &  \\
2016-05-27 & 16:29:34 & 135 & 3~GHz & 0 &  &  \\
2016-08-23 & 17:26:28 & 54 & 3~GHz & 1 &  &  \\
2016-09-02 & 15:52:17 & 54 & 3~GHz & 2 &  &  \\
2016-09-07 & 10:14:50 & 120 & 3~GHz & 1 &  &  \\
2016-09-08 & 10:14:40 & 120 & 3~GHz & 0 &  &  \\
2016-09-10 & 9:59:36 & 120 & 3~GHz & 0 &  &  \\
2016-09-11 & 9:59:49 & 120 & 3~GHz & 0 &  &  \\
2016-09-12 & 9:15:19 & 120 & 3~GHz & 1 & AMI-LA &  \\
2016-09-13 & 9:23:59 & 120 & 3~GHz & 0 &  &  \\
2016-09-14 & 9:20:23 & 120 & 3~GHz & 1 & Arecibo, AMI-LA &  \\
2016-09-15 & 9:16:29 & 120 & 3~GHz & 1 &  &  \\
2016-09-16 & 9:11:23 & 120 & 3~GHz & 0 &  &  \\
2016-09-17 & 8:59:20 & 120 & 3~GHz & 1 & Arecibo, Effelsberg, AMI-LA & Arecibo \\
2016-09-18 & 8:59:27 & 120 & 3~GHz & 1 & Arecibo, Effelsberg, AMI-LA &  \\
2016-09-19 & 8:44:32 & 120 & 3~GHz & 0 &  &  \\
2016-09-20 & 8:44:33 & 120 & 3~GHz & 0 &  &  \\
2016-09-20 & 17:19:3 & 120 & 6~GHz & 0 &  &  \\
2016-09-21 & 9:13:57 & 120 & 6~GHz & 0 &  &  \\
2016-09-22 & 9:12:24 & 120 & 6~GHz & 0 &  &  \\
\enddata
\tablecomments{Arecibo, Effelsberg, and AMI-LA observations were conducted at 1.4, 4.5, and 15~GHz, respectively.}
%\end{tabular}
%\tablenotetext{a}{}
%\label{tab:obs}
\end{deluxetable*}

\bibliographystyle{yahapj}

\section*{Acknowledgements}
% observatories
The National Radio Astronomy Observatory is a facility of the National Science Foundation operated under cooperative agreement by Associated Universities, Inc..
Part of this research was carried out at the Jet Propulsion Laboratory, California Institute of Technology, under a contract with the National Aeronautics and Space Administration.
% systems
This research made use of Astropy, a community-developed core Python package for Astronomy (Astropy Collaboration, 2013).
%individuals
CJL is supported by the University of California Office of the President under Lab Fees Research Program Award 237863 and NSF award 1611606. MAM is supported by NSF award \#1458952. KPM's research is supported by the Oxford Centre for Astrophysical Surveys which is funded through the Hintze Family Charitable Foundation. AS gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE. The AMI-LA telescope gratefully acknowledges support from the European Research Council under grant ERC-2012- StG-307215 LODESTONE, the UK Science and Technology Facilities Council (STFC) and the University of Cambridge. PS holds a Covington Fellowship at DRAO.
VMK acknowledges support from NSERC, CIFAR, the Canada Research Chair Program and from the Lorne Trottier Chair in Astrophysics \& Cosmology.

We thank Liam Connor for useful feedback and the VLA staff for their support of \rf\ development. We acknowledge partial support from the Research Corporation for Scientific Advancement (RCSA) for participation in the meeting Fast Radio Bursts: New Probes of Fundamental Physics and Cosmology at the Aspen Center for Physics (February 12-17, 2017).

\bibliography{fasttrants.bib}

\end{document}